PyQBench: a Python library for benchmarking gate-based quantum computers

### 2023-02-28 txt/manuscript.tex
add ocean

\begin{table}[H]
\begin{tabular}{|l|p{5.5cm}|p{7.5cm}|}
\hline
C1 & Current code version & 0.1.1 \\
\hline
C2 & Permanent link to code/repository used for this code version & \url{https://github.com/iitis/PyQBench} \\
\hline
C3 & Code Ocean compute capsule & \todo[inline]{???}\\
\hline
C4 & Legal Code License & Apache License 2.0\\
\hline
C5 & Code versioning system used & git \\
\hline
C6 & Software code languages, tools, and services used & Python, Qiskit, AWS Braket \\
\hline
C7 & Compilation requirements, operating environments \& dependencies &
\texttt{Python >= 3.8}\newline
\texttt{numpy \textapprox= 1.22.0}\newline
\texttt{scipy \textapprox= 1.7.0}\newline
\texttt{pandas \textapprox= 1.5.0}\newline
\texttt{amazon-braket-sdk >= 1.11.1}\newline
\texttt{pydantic \textapprox= 1.9.1}\newline
\texttt{qiskit \textapprox= 0.37.2}\newline
\texttt{mthree \textapprox= 1.1.0}\newline
\texttt{tqdm \textapprox= 4.64.1}\newline
\texttt{pyyaml \textapprox= 6.0}\newline
\texttt{qiskit-braket-provider \textapprox= 0.0.3}\\
\hline
C8 & If available Link to developer documentation/manual &
\url{https://pyqbench.readthedocs.io/en/latest/}\\
\hline
C9 & Support email for questions & \url{dexter2206@gmail.com}\\
\hline
\end{tabular}
\caption{Code metadata}
\label{}
\end{table}

@@

It is well known that NISQ devices have their limitations \cite{preskillnew}. The question is to what extent those
devices can perform meaningful computations? To answer this question, one has to devise a
methodology for benchmarking them. For gate--based computers, on which this paper focuses, there
already exist several approaches.
One could mention randomized benchmarking \cite{liu2022sampling, knill2007randomized, knill2008randomized, wallman2014randomized, helsen2022general,
	cornelissen2021scalable}, benchmarks based on the quantum volume \cite{cross2019validating, moll2018quantum, pelofske2022volume} or cross-entropy benchmarking \cite{arute2019quantum, neill2018blueprint}.

@@

\todo[inline]{cos tu dopisac o tych metodach}
Another quantity used for benchmarking NISQ devices is quantum volume. The quantum volume
characterizes capacity of a device for solving computational problems. It takes into account
multiple factors like number of qubits, connectivity and measurement errors. The Qiskit library allows one to
measure quantum volume of a device by using its \texttt{qiskit.ignis.verification.quantum\_volume}.
Other implementations of Quantum Volume can be found as well, see e.g. \cite{volume-in-practice}.

@@

\item Preparing configuration files describing the backend and the experiment scenario.
\item Submitting/running experiments. Depending on the experiment scenario, execution can be synchronous, or asynchronous.
\item (optional) Checking the status of the submitted jobs if the execution is asynchronous.
\item Resolving asynchronous jobs into the actual measurement outcomes.
\item Converting obtained measurement outcomes into tabulated form.
\end{enumerate}
\subsubsection{Preparing configuration files}
The configuration of PyQBench CLI is driven by YAML files. The first configuration file
describes the experiment scenario to be executed. The second file describes the backend.
Typically, this backend will correspond to the physical device to be benchmarked, but for testing
purposes one might as well use any other Qiskit--compatible backend including simulators.
Let us first describe the experiment configuration file, which might look as follow.
\begin{lstlisting}[language=Python, caption=Defining the experiment, label=lst:experiment]
type: discrimination-fourier
qubits:
	- target: 0
	  ancilla: 1
	- target: 1
      ancilla: 2
angles:
	start: 0
	stop: 2 * pi
	num_steps: 3
gateset: ibmq
method: direct_sum
num_shots: 100
\end{lstlisting}
The experiment file contains the following fields:
\begin{itemize}
	\item \texttt{type}: a string describing the type of the experiment. Currently, the only option of  \textit{type} is \texttt{discrimination-fourier}.
	\item \texttt{qubits}: a list enumerating pairs of qubits on which the experiment should be run.
	For configuration in listing \ref{lst:experiment}, the benchmark will run on two pairs of qubits. The first pair is 0 and 1, and
	the second one is 1 and 2. We decided to describe a pair by using \texttt{target} and
	\texttt{ancilla} keys rather than using a plain list to emphasize that the role of qubits in the
	experiment is not symmetric.
	\item \texttt{angles}: an object describing the range of angles for Fourier parameterized
	family. The described range is always uniform, starts at the \texttt{start}, ends at
	\texttt{stop} and contains \texttt{num\_steps} points, including both \texttt{start} and
	\texttt{stop}. The \texttt{start} and \texttt{stop} can be arithmetic expressions using
	\texttt{pi} literal. For instance, the range defined in listing \ref{lst:experiment} contains
	three points: 0, $\pi$ and $2\pi$.
 	 \item \texttt{gateset}: a string describing the set of gates used in the decomposition of
 	 circuits in the experiment. The PyQBench contains explicit implementations of circuits The
 	 possible options are \textit{[ibmq, lucy, rigetti]}, corresponding to decompositions
 	 compatible with IBM Q devices, OQC Lucy device, and Rigetti devices. Alternatively, one might
 	 wish to turn off the decomposition by using a special value \texttt{generic}. However, for this
 	 to work a backend used for the experiment must natively implement all the gates needed for the
 	 experiment, as described in \ref{sec:fourier}.
 	\item \texttt{method}: a string, either \texttt{postselection} or \texttt{direct\_sum}
 	determining which implementation of the conditional measurement is used.
 	\item \texttt{num\_shots}: an integer defines how many shots are performed in the experiment for
 	a particular angle, qubit pair and circuit. Note that if one wishes to compute the total number
 	of shots in the experiment, it is necessary to take into account that the \texttt{postselection}
 	method uses twice as many circuits as the \texttt{direct\_sum} method.
\end{itemize}

@@

For synchronous experiment, the part of output looks similar to the one below. The whole YAML file can be seen in \ref{app:example}.
%Consider the particular experiment data
\begin{lstlisting}[language=Python]
data:
- target: 0
  ancilla: 1
  phi: 0.0
  results_per_circuit:
  - name: id
  histogram: {'00': 28, '01': 26, '10': 21, '11': 25}
  mitigation_info:
	target: {prob_meas0_prep1: 0.052200000000000024, prob_meas1_prep0: 0.0172}
	ancilla: {prob_meas0_prep1: 0.05900000000000005, prob_meas1_prep0: 0.0202}
  mitigated_histogram: {'00': 0.2637212373658018, '01': 0.25865061319892463, '10': 0.2067279352110304, '11': 0.2709002142242433}
\end{lstlisting}
	The data includes \textit{target, ancilla, phi},
	and \textit{results\_per\_circuit}.   The first three pieces of information have already been described. The last data  \textit{results\_per\_circuit} gives us
	the following additional information:
	\begin{itemize}
		\item  \texttt{name}:  the information which measurement is used during experiment, either
		string \texttt{"u"} for $\PP_{U}$ or string \texttt{"id"} for $\PP_{\Id}$. In this example
		we consider $\PP_{\Id}$.
		\item \texttt{histogram}: the dictionary with measurement's outcomes. The keys represent possible bitstrings, whereas the values present number of occurrences.
		\item \texttt{mitigation\_info}:  \textit{system.properties().qubits} contains information on the physical attributes of the qubits. This includes e.g. information about readout error. As shown below, the \textit{readout\_error} is further partitioned into \textit{prob\_meas0\_prep1} and \textit{prob\_meas1\_prep0}, indicating the probability of preparing a given computational basis state, but measuring the orthogonal state.  The readout error is taken to be the average of these two values. Note that \textit{prob\_meas0\_prep1} is usually the dominant factor in the readout error due to relaxation during the measurement process.  ~\cite{mthree, mthreepublication}
		\item mitigated\_histogram:  the histogram with measurement's outcomes after mitigation.
	\end{itemize}

@@

\subsubsection{Resolving asynchronous jobs}
If the status of jobs is \textit{DONE}, we have to obtain measurements from the submitted jobs. This is done using the following command:
\begin{lstlisting}[language=Python]
qbench disc-fourier resolve async-results.yml resolved.yml
\end{lstlisting}
The resolved results, stored in \textit{resolved.yml}, would look just like the experiment was run synchronously. More precisely, we should have a file containing measurement histograms, no matter if we run a synchronous or asynchronous experiment (see \ref{app:example}). Next, we can compute the discrimination probabilities.

\subsubsection{Computing probabilities}
Regardless of whether we use a synchronous or asynchronous approach, we can convert the resulting YML file \textit{results.yml} to a CSV file by creating a result table in which we place the results of discrimination probabilities.
To construct a table with probabilities, you can use the following command
\begin{lstlisting}[language=Python]
qbench disc-fourier tabulate results.yml results.csv
\end{lstlisting}
A single row for each tuple of \textit{(target, ancilla, phi)}. The \textit{disc\_prob} column will contain the discrimination probability for given configuration. The \textit{mit\_disc\_prob} column contains discrimination probability if Mthree mitigation was applied~\cite{mthree, mthreepublication}.
A sample csv file is provided below:
\begin{figure}[h!]
	\centering
	\includegraphics[scale=1]{results/async_result}
	\caption{Tabulate results}
	\label{fig:tabulateresults}
\end{figure}

@@

\bibitem{arute2019quantum} Arute, F., Arya, K., Babbush, R., Bacon, D., Bardin, J. C., Barends, R., ... Martinis, J. M. (2019). Quantum supremacy using a programmable superconducting processor. Nature, 574(7779), 505-510.
\bibitem{neill2018blueprint}
Neill, C., Roushan, P., Kechedzhi, K., Boixo, S., Isakov, S. V., Smelyanskiy, V., ...  Martinis, J. M. (2018). A blueprint for demonstrating quantum supremacy with superconducting qubits. Science, 360(6385), 195-199.

@@

For the same experiment file, we use the flag \textit{asynchronous: true} to define asynchronous experiment.

>>>>

\begin{table}[H]
\begin{tabular}{|l|p{5.5cm}|p{7.5cm}|}
\hline
C1 & Current code version & 0.1.1 \\
\hline
C2 & Permanent link to code/repository used for this code version & \url{https://github.com/iitis/PyQBench} \\
\hline
C3 & Code Ocean compute capsule & \texttt{https://codeocean.com/capsule/ 89088992-9a27-4712-8525-} \texttt{d92a9b23060f/tree}\\
\hline
C4 & Legal Code License & Apache License 2.0\\
\hline
C5 & Code versioning system used & git \\
\hline
C6 & Software code languages, tools, and services used & Python, Qiskit, AWS Braket \\
\hline
C7 & Compilation requirements, operating environments \& dependencies &
\texttt{Python >= 3.8}\newline
\texttt{numpy \textapprox= 1.22.0}\newline
\texttt{scipy \textapprox= 1.7.0}\newline
\texttt{pandas \textapprox= 1.5.0}\newline
\texttt{amazon-braket-sdk >= 1.11.1}\newline
\texttt{pydantic \textapprox= 1.9.1}\newline
\texttt{qiskit \textapprox= 0.37.2}\newline
\texttt{mthree \textapprox= 1.1.0}\newline
\texttt{tqdm \textapprox= 4.64.1}\newline
\texttt{pyyaml \textapprox= 6.0}\newline
\texttt{qiskit-braket-provider \textapprox= 0.0.3}\\
\hline
C8 & If available Link to developer documentation/manual &
\url{https://pyqbench.readthedocs.io/en/latest/}\\
\hline
C9 & Support email for questions & \url{dexter2206@gmail.com}\\
\hline
\end{tabular}
\caption{Code metadata}
\label{}
\end{table}

@@

It is well known that NISQ devices have their limitations \cite{preskillnew}. The question is to what extent those
devices can perform meaningful computations? To answer this question, one has to devise a
methodology for benchmarking them. For gate--based computers, on which this paper focuses, there
already exist several approaches.
One could mention randomized benchmarking \cite{liu2022sampling, knill2007randomized, knill2008randomized, wallman2014randomized, helsen2022general,
	cornelissen2021scalable}, benchmarks based on the quantum volume \cite{cross2019validating, moll2018quantum, pelofske2022volume}. 
%or cross-entropy benchmarking \cite{arute2019quantum, neill2018blueprint}.

@@

Another quantity used for benchmarking NISQ devices is quantum volume. The quantum volume
characterizes capacity of a device for solving computational problems. It takes into account
multiple factors like number of qubits, connectivity and measurement errors. The Qiskit library allows one to
measure quantum volume of a device by using its \texttt{qiskit.ignis.verifica \\ tion.quantum\_volume}.
Other implementations of Quantum Volume can be found as well, see e.g. \cite{volume-in-practice}.

@@

\item Preparing configuration files describing the backend and the experiment scenario.
\item Submitting/running experiments. Depending on the experiment scenario, execution can be synchronous, or asynchronous.
\item (optional) Checking the status of the submitted jobs if the execution is asynchronous.
\item Resolving asynchronous jobs into the actual measurement outcomes.
\item Converting obtained measurement outcomes into tabulated form.
\end{enumerate}
\subsubsection{Preparing configuration files}
The configuration of PyQBench CLI is driven by YAML files. The first configuration file
describes the experiment scenario to be executed. The second file describes the backend.
Typically, this backend will correspond to the physical device to be benchmarked, but for testing
purposes one might as well use any other Qiskit--compatible backend including simulators.
Let us first describe the experiment configuration file, which might look as follow.
\begin{lstlisting}[language=Python, caption=Defining the experiment, label=lst:experiment]
type: discrimination-fourier
qubits:
	- target: 0
	  ancilla: 1
	- target: 1
      ancilla: 2
angles:
	start: 0
	stop: 2 * pi
	num_steps: 3
gateset: ibmq
method: direct_sum
num_shots: 100
\end{lstlisting}
The experiment file contains the following fields:
\begin{itemize}
	\item \texttt{type}: a string describing the type of the experiment. Currently, the only option of  \texttt{type} is \texttt{discrimination-fourier}.
	\item \texttt{qubits}: a list enumerating pairs of qubits on which the experiment should be run.
	For configuration in listing \ref{lst:experiment}, the benchmark will run on two pairs of qubits. The first pair is 0 and 1, and
	the second one is 1 and 2. We decided to describe a pair by using \texttt{target} and
	\texttt{ancilla} keys rather than using a plain list to emphasize that the role of qubits in the
	experiment is not symmetric.
	\item \texttt{angles}: an object describing the range of angles for Fourier parameterized
	family. The described range is always uniform, starts at the \texttt{start}, ends at
	\texttt{stop} and contains \texttt{num\_steps} points, including both \texttt{start} and
	\texttt{stop}. The \texttt{start} and \texttt{stop} can be arithmetic expressions using
	\texttt{pi} literal. For instance, the range defined in listing \ref{lst:experiment} contains
	three points: 0, $\pi$ and $2\pi$.
 	 \item \texttt{gateset}: a string describing the set of gates used in the decomposition of
 	 circuits in the experiment. The PyQBench contains explicit implementations of circuits The
 	 possible options are \texttt{[ibmq, lucy, rigetti]}, corresponding to decompositions
 	 compatible with IBM Q devices, OQC Lucy device, and Rigetti devices. Alternatively, one might
 	 wish to turn off the decomposition by using a special value \texttt{generic}. However, for this
 	 to work a backend used for the experiment must natively implement all the gates needed for the
 	 experiment, as described in \ref{sec:fourier}.
 	\item \texttt{method}: a string, either \texttt{postselection} or \texttt{direct\_sum}
 	determining which implementation of the conditional measurement is used.
 	\item \texttt{num\_shots}: an integer defines how many shots are performed in the experiment for
 	a particular angle, qubit pair and circuit. Note that if one wishes to compute the total number
 	of shots in the experiment, it is necessary to take into account that the \texttt{postselection}
 	method uses twice as many circuits as the \texttt{direct\_sum} method.
\end{itemize}

@@

For synchronous experiment, the part of output looks similar to the one below. The whole YAML file can be seen in \ref{app:example}.
%Consider the particular experiment data
\begin{lstlisting}[language=Python]
data:
- target: 0
  ancilla: 1
  phi: 0.0
  results_per_circuit:
  - name: id
  histogram: {'00': 28, '01': 26, '10': 21, '11': 25}
  mitigation_info:
	target: {prob_meas0_prep1: 0.052200000000000024, prob_meas1_prep0: 0.0172}
	ancilla: {prob_meas0_prep1: 0.05900000000000005, prob_meas1_prep0: 0.0202}
  mitigated_histogram: {'00': 0.2637212373658018, '01': 0.25865061319892463, '10': 0.2067279352110304, '11': 0.2709002142242433}
\end{lstlisting}
	The data includes \texttt{target, ancilla, phi},
	and \texttt{results\_per\_circuit}.   The first three pieces of information have already been described. The last data  \texttt{results\_per\_circuit} gives us
	the following additional information:
	\begin{itemize}
		\item  \texttt{name}:  the information which measurement is used during experiment, either
		string \texttt{"u"} for $\PP_{U}$ or string \texttt{"id"} for $\PP_{\Id}$. In this example
		we consider $\PP_{\Id}$.
		\item \texttt{histogram}: the dictionary with measurement's outcomes. The keys represent possible bitstrings, whereas the values present number of occurrences.
		\item \texttt{mitigation\_info}:  \texttt{system.properties().qubits} contains information on the physical attributes of the qubits. This includes e.g. information about readout error. As shown below, the \texttt{readout\_error} is further partitioned into \texttt{prob\_meas0\_prep1} and \texttt{prob\_meas1\_prep0}, indicating the probability of preparing a given computational basis state, but measuring the orthogonal state.  The readout error is taken to be the average of these two values. Note that \texttt{prob\_meas0\_prep1} is usually the dominant factor in the readout error due to relaxation during the measurement process.  ~\cite{mthree, mthreepublication}
		\item mitigated\_histogram:  the histogram with measurement's outcomes after mitigation.
	\end{itemize}

@@

\subsubsection{Resolving asynchronous jobs}
If the status of jobs is \texttt{DONE}, we have to obtain measurements from the submitted jobs. This is done using the following command:
\begin{lstlisting}[language=Python]
qbench disc-fourier resolve async-results.yml resolved.yml
\end{lstlisting}
The resolved results, stored in \texttt{resolved.yml}, would look just like the experiment was run synchronously. More precisely, we should have a file containing measurement histograms, no matter if we run a synchronous or asynchronous experiment (see \ref{app:example}). Next, we can compute the discrimination probabilities.

\subsubsection{Computing probabilities}
Regardless of whether we use a synchronous or asynchronous approach, we can convert the resulting YML file \texttt{results.yml} to a CSV file by creating a result table in which we place the results of discrimination probabilities.
To construct a table with probabilities, you can use the following command
\begin{lstlisting}[language=Python]
qbench disc-fourier tabulate results.yml results.csv
\end{lstlisting}
A single row for each tuple of \texttt{(target, ancilla, phi)}.  The \texttt{ideal\_prob} 
column says the theoretical probability of correct discrimination whereas
the \texttt{disc\_prob} column contains the empirical discrimination probability. The \texttt{mit\_disc\_prob} column contains discrimination probability if Mthree mitigation was applied~\cite{mthree, mthreepublication}.
A sample csv file is provided below:
\begin{figure}[h!]
	\centering
	\includegraphics[scale=1]{results/async_result}
	\caption{Tabulate results}
	\label{fig:tabulateresults}
\end{figure}

@@

%\bibitem{arute2019quantum} Arute, F., Arya, K., Babbush, R., Bacon, D., Bardin, J. C., Barends, R., ... Martinis, J. M. (2019). Quantum supremacy using a programmable superconducting processor. Nature, 574(7779), 505-510.
%\bibitem{neill2018blueprint}
%Neill, C., Roushan, P., Kechedzhi, K., Boixo, S., Isakov, S. V., Smelyanskiy, V., ...  Martinis, J. M. (2018). A blueprint for demonstrating quantum supremacy with superconducting qubits. Science, 360(6385), 195-199.

@@

For the same experiment file, we use the flag \texttt{asynchronous: true} to define asynchronous experiment.
